Formal Verification of Real Models from Platforms Using The B-method

### 2009-04-27 papers/FM2009/source/paper.tex

[Conclusao]This project has many proof obligations and its process of verification is the step of the most time
consumed. The verification step of all the model spent about a month. This was hard, because was necessary to build,
correct and prove several expressions from basic projects, Hardware and Types Library. However, when the basic
projects are ready, the modelling and verification of new plataform \footnote{Actualy, the microprocessors M8051 and
PIC16F432 are under construction} become more fast and easy. Thus, these basic projects were also built to re-use
to other platforms.

When the basic projects are ready, then about 89\% proof non obvious and 55\% of Proofs WD was conclused
automaticly. The others were concluded through the use of commands ...
Com os projetos de base completamente provados, então aproximadamente 89\% das provas não obvias e 55\% da
Proofs WD foram realizadas automaticamente. As demais provas foram realizadas após o uso de alguns passos
de prova\footnote{Passos de prova são comandos que indicam um caminho para o provador concluir a prova,
esses comandos apenas indicam um caminho, ou seja, ele não introduzem hipósteses falsas} associados a
pattern matching of proof obligations. The user pass associeted to match pattern  aceleraram bastante o
processo de prova, pois existem muitas instruções do microcontrolador com caractéristicas semelhantes e o
atelierb é capaz de re-usar os passos de prova. Então, poucos passos de prova podem ser úteis para uma
grande variedade de provas. Um exemplo é o conjunto de 17 passos de prova que rapidamente ajudou a
verificar 99\% (2295) das provas WD. A experiência em adicionar passos de prova também ajudou a reduzir o
tempo do processo de prova em mais de 90\%. Isso aconteceu, porque o uso de passos de prova direciona melhor o
provador ao fim da prova.

To further reduce cost of proof process, a new feature supported by AtelierB was used, the
parallelization. To do this were used three computers each one with two cores, the that theoretically
increased by six times the capacity of processing.

A special care was taken to represent the model in way simple and accurate. This care also reduced the number of
proofs. Because, the simple act of optimizing the use the simple act of optimizing the use of
override operator ($\lover$) fell too the number of proof obligations.

@@

 \section{Conlusitions}
[I need review the english]
% [intro]
A necessidade comercial para produzir \textit{hardware} e \textit{software} de alta qualidade é sempre
crescente. O métodos B já têm demonstrado sucesso na especificação de sistemas comerciais críticos.

>>>>

% [Conclusao]
This project has many proof obligations and its process of verification is the step of the most time consumed. The
verification step of all the model spent about a month. This was hard, because was necessary to build, correct and
prove several expressions from basic projects, Hardware and Types Library. However, when the basic projects are
ready, the modelling and verification of new plataform \footnote{Actualy, the microprocessors M8051 and PIC16F432
are under construction} become more fast and easy. Thus, these basic projects were also built to re-use to other
platforms. When the basic projects are ready, then about 89\% proof non obvious and 55\% of Proofs WD was conclused
automaticly.

The others were concluded through the use of proof commands\footnote{The proof commands are step that just indicate
a walk to prover make the proof, thus, theses commands cannot introduce false hypotheses.} associated to match
pattern from proof obligations. The match pattern accelerates too much the prove process, because there are many
similar assembler instructions and the AtelierB can re-use the match pattern and its proof commands associated.
Therefore, few proof commands can be useful to many obrigações de prova. A good example is a set of 17 proof
commands that fastly helps to verify 99\% (2295) of proofs WD. To further reduce cost of proof process, a new feature
supported by AtelierB was used, the parallelization. To do this were used three computers each one with two cores,
the that theoretically increased by six times the capacity of processing.

A special care was taken to represent the model in way simple and accurate. This care also reduced the number of
proofs. Because, the simple act of optimizing the use of override operator ($\lover$) fell too the number of proof
obligations.

@@

 \section{Related Works}

% This work had as the main motivation the  verififcation in assembler level and consequently the
% verification of harware model. The model created express the semantic of functioning of hardware and
% with this model is possible verify the semantic of hardware components.

There are in the literature of computer science some technics\cite{BHDL_2003,GEMPLUS_99} to model hardware
and the virtual machines using the B method. Then, the B method has been used successfully to model the
operational semantic, however The cost modeling was still too expensive and this paper describes some
techniques to lower the cost modeling. In general, the researchers are concetrated at model the Java
Virtual Machine (JVM), but it is not easy to specify  it, because it very complex, and become more
difficult to write good lemmas, to normalize the functions and to verify the properties. Then, this
technique specify a model more simple to get better results. Besides, the objectives are to get concret
results fastly and, in the same time, stay near to model of actual processors.
% Re-uso  - modularization
The most of related works are specific to one plataform and its basic definitions cannot be re-used at in
other platforms. The used strategy created a good logic separation about the concepts of modelling. Thus,
the basic projects, the Hardware and Types Library, have functions with a low coupling and high
cohesion, that promove the re-use. Soon, these re-use better too the code generation and prove
process.{{Pare aqui \ssssssss }}

 [Uma das grandes vantagens é a evolução suporte ao método B]

 A other advantage of this work is the use of news and advanceds funcionalities of B method. The first is
 the paralelization of prove procces and  the second is instensive use of math patterns associated with proof
 commands. In the

 [AMM]
Verificação formal já vem sendo aplicada com sucesso na verificação \ldots Porém os custos eram altos,
pois se passou X tempo para modelar \ldots e

A modelagem do microcontrolador foi construida por um unico designer and consumiu aproximadamente six
month to study the plattaform Z80, model, adjust to simplifi and prove.

 \section{Conlusitions}

 % Trab. Futuros
 %- Refinar o modelo para gerar através dos  código e apartir deste código criar um
 %interpretador/simulador
 %- Construir compilador

%  \begin{itemize}
%    Citar ambuiguidades encontradas no manual e eventuais erros encontrados.
%   \item Utilidade o processo de desenvolvimento do hardware:
%   \subitem Ajuda na remoção de eventuais ambiguidades providas de uma descrição informal.
%   \subitem Verificação de propriedades: sobre a manipulação e tipo dos dados.
%   \subitem Documentação formalizada matemáticamente
%   \subitem Possibilidade da geracao de código VHDL apartir do modelo B.
%   \subitem Colaboracao para construção de um simulador/interpretador
% \end{itemize}

[I need review the english]
% [intro]
A necessidade comercial para produzir \textit{hardware} e \textit{software} de alta qualidade é sempre
crescente. O métodos B já têm demonstrado sucesso na especificação de sistemas comerciais críticos.
